---
output:
  pdf_document: default
  html_document: default
---
Vectorising a portion of the global forest cover map
========================================================

The data can be obtained from here.
http://earthenginepartners.appspot.com/science-2013-global-forest


Load the packages.

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(rgdal)
library(rgeos)
library(raster)
library(dismo)

```


```{r,eval=FALSE}
## These lines are for troubleshooting if a student has a problem
## Not run normally
system ("gdalinfo forest2000.tif")
system ("gdalinfo loss.tif")
```

These lines should reproject if the layer is in the wrong CRS.

```{r,eval=TRUE}
system ("gdalwarp -t_srs EPSG:3857 forest2000.tif forest2000_3857.tif")
system ("gdalwarp -t_srs EPSG:3857 loss.tif loss_3857.tif")
```



## Load the data

```{r}

forest2000<-raster("forest2000_3857.tif")
loss<-raster("loss_3857.tif")
mp<-gmap(forest2000,type="satellite")
plot(mp)
axis(1)
axis(2)
box()
grid()

```

## Turn into binary forest-non forest maps

You may want to change the cut off point here. I have set it to 80%

```{r}
percent_cover= 80
```

This is good for moist forest, but you may need a lower value for dry forest.

```{r}

forest2000<-1*(forest2000>=percent_cover)
nonforest2000<-1-forest2000
mask<-1-loss
forest2012<-forest2000*mask
nonforest2012<-1-forest2012

```


## Set the minimum size for the fragments.

### You can also change this, but don't set it too small

The size is in hectares
One hectare is around 9 pixels in the original image

```{r}

min_size<-4

```

## PLot the layers

```{r,eval=T}

plot(forest2000)
plot(nonforest2000)
plot(forest2012)
plot(nonforest2012)

```

## Calculate the total area in hectares

```{r}

totalarea<-(forest2000@extent@xmax-forest2000@extent@xmin)*(forest2000@extent@ymax-forest2000@extent@ymin)/10000
totalarea

```
## Vectorisation function


```{r,message=FALSE,warning=FALSE}
vect<-function(){
  id <- values(cl)
crds <- coordinates(cl)
crds <- crds[!is.na(id), ]
id <- id[!is.na(id)]
crds <- data.frame(id, crds)
coordinates(crds) <- ~x + y
buf1 <- gBuffer(crds, width = 50, byid = T)
buf2 <- gUnaryUnion(buf1, id)
buf3 <- gBuffer(buf2, width = -50, byid = T)
frags<-buf3[gArea(buf3,byid=T)>min_size*10000]
area<-gArea(frags,byid=T)
edge<-gBoundary(frags,byid=T)
perims<-gLength(edge,byid=T)
d<-data.frame(id=names(frags),area,perims)
d$shape<-d$perims/(2*pi*sqrt(d$area/pi))
d$area<-d$area/10000
d<-d[order(-d$area),]
d$cumarea<-cumsum(d$area)
d$ptot<-d$area/totalarea*100
d$parea<-d$area/sum(d$area)*100
frags<-SpatialPolygonsDataFrame(frags,data=d,match.ID=FALSE)
proj4string(frags)<-CRS("+init=epsg:3857")
frags
  
}



```

## Vectorise forest 2000 area

```{r,message=FALSE,warning=FALSE}
cl<-clump(forest2000)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()

writeOGR(frags,dsn="shapefiles","forest2000",driver="ESRI Shapefile",over=TRUE)
```

## Tabular data 

The table includes the percentage of the total study site area for each fragment and the percentage of the landscape class. This is particularly useful when looking at the largest patch.

```{r}
frags@data
```


```{r}
plot(mp)
plot(frags, add = T, border = "red", col = "transparent")

```

## Vectorise forest 2012



```{r,message=FALSE,warning=FALSE}

cl<-clump(forest2012)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()
frags@data
writeOGR(frags,dsn="shapefiles","forest2012",driver="ESRI Shapefile",over=TRUE)

```

```{r}
plot(mp)
plot(frags, add = T, border = "red", col = "transparent")
```

## Vectorise non forest 2000

```{r,message=FALSE,warning=FALSE}
cl<-clump(nonforest2000)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()
frags@data
writeOGR(frags,dsn="shapefiles","nonforest2000",driver="ESRI Shapefile",over=TRUE)

```

## Vectorise non forest 2012

```{r,message=FALSE,warning=FALSE}

cl<-clump(nonforest2012)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()
frags@data
writeOGR(frags,dsn="shapefiles","nonforest2012",driver="ESRI Shapefile",over=TRUE)

```


