---
output:
  pdf_document: default
  html_document: default
---
Vectorising a portion of the global forest cover map
========================================================

The data can be obtained from here.
http://earthenginepartners.appspot.com/science-2013-global-forest


Load the packages.

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(rgdal)
library(rgeos)
library(raster)
library(dismo)

```


```{r,eval=FALSE}
system ("gdalinfo forest2000.tif")
system ("gdalinfo loss.tif")
```

If you see that your layer is in fact in EPSG then you need to run the lines below.

```{r,eval=TRUE}
system ("gdalwarp -t_srs EPSG:3857 forest2000.tif forest2000_3857.tif")
system ("gdalwarp -t_srs EPSG:3857 loss.tif loss_3857.tif")
```



## Load the data

```{r}

forest2000<-raster("forest2000_3857.tif")
loss<-raster("loss_3857.tif")
mp<-gmap(forest2000,type="satellite")
plot(mp)
axis(1)
axis(2)
box()
grid()

```

## Turn into binary forest-non forest maps

You can change the cut off point here. I have set it to 50%

```{r}
percent_cover= 80

forest2000<-1*(forest2000>=percent_cover)
nonforest2000<-1-forest2000
mask<-1-loss
forest2012<-forest2000*mask
nonforest2012<-1-forest2012

```


## Set the minimum size for the fragments.

### CHANGE THIS IF YOU WANT, BUT BE CAREFUL NOT TO MAKE TOO MANY SMALL AREAS. 

The size is in hectares
One hectare is around 9 pixels in the original image

```{r}

min_size<-4

```


```{r,eval=T}

plot(forest2000)
plot(nonforest2000)
plot(forest2012)
plot(nonforest2012)

```



```{r}

totalarea<-(forest2000@extent@xmax-forest2000@extent@xmin)*(forest2000@extent@ymax-forest2000@extent@ymin)/10000
totalarea

```
## Vectorisation function


```{r,message=FALSE,warning=FALSE}
vect<-function(){
  id <- values(cl)
crds <- coordinates(cl)
crds <- crds[!is.na(id), ]
id <- id[!is.na(id)]
crds <- data.frame(id, crds)
coordinates(crds) <- ~x + y
buf1 <- gBuffer(crds, width = 50, byid = T)
buf2 <- gUnaryUnion(buf1, id)
buf3 <- gBuffer(buf2, width = -50, byid = T)
frags<-buf3[gArea(buf3,byid=T)>min_size*10000]
area<-gArea(frags,byid=T)
edge<-gBoundary(frags,byid=T)
perims<-gLength(edge,byid=T)
d<-data.frame(id=names(frags),area,perims)
d$shape<-d$perims/(2*pi*sqrt(d$area/pi))
d$area<-d$area/10000
d<-d[order(-d$area),]
d$cumarea<-cumsum(d$area)
d$ptot<-d$area/totalarea*100
d$parea<-d$area/sum(d$area)*100
frags<-SpatialPolygonsDataFrame(frags,data=d,match.ID=FALSE)
proj4string(frags)<-CRS("+init=epsg:3857")
frags
  
}



```

## Vectorise forest 2000 area

```{r,message=FALSE,warning=FALSE}
cl<-clump(forest2000)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()
frags@data
writeOGR(frags,dsn="shapefiles","forest2000",driver="ESRI Shapefile",over=TRUE)
```


```{r}
plot(mp)
plot(frags, add = T, border = "red", col = "transparent")

```

## Vectorise forest 2012



```{r,message=FALSE,warning=FALSE}

cl<-clump(forest2012)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()
frags@data
writeOGR(frags,dsn="shapefiles","forest2012",driver="ESRI Shapefile",over=TRUE)

```

```{r}
plot(mp)
plot(frags, add = T, border = "red", col = "transparent")
```

## Vectorise non forest 2000

```{r,message=FALSE,warning=FALSE}
cl<-clump(nonforest2000)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()
frags@data
writeOGR(frags,dsn="shapefiles","nonforest2000",driver="ESRI Shapefile",over=TRUE)

```

## Vectorise non forest 2012

```{r,message=FALSE,warning=FALSE}

cl<-clump(nonforest2012)
frags<-vect()
plot(frags,col="darkgreen")
axis(1)
axis(2)
box()
grid()
frags@data
writeOGR(frags,dsn="shapefiles","nonforest2012",driver="ESRI Shapefile",over=TRUE)

```


